---
title: "Clustering players"
output: html_notebook
---

Load packages

```{r}
library(tidyverse)
#also uses trimcluster 0.1-5 for trimmed clustering
#also uses fpc 2.2-13 to derive statistics of clustering goodness
```
# Data load and prep

Load files created in 'data_prep.R'

```{r}
games_table <- read.csv(".\\data\\luckybit_games_table.csv")
luckybit_addrIDs_raw <- read.csv(".\\data\\luckybit_games_address.csv")
bets_complete <- read.csv(".\\data\\luckybit_bets_usered.csv")
```

Add payout variations (CV or coefficient of variation) to the games

```{r}
luckybit_addrIDs <- 
  games_table %>%
  pivot_longer(c("blue","green","yellow","red"), names_to = "Name", values_to = "multiplier") %>%
  group_by(Name) %>%
  mutate(exp_ret = sum(p_win*multiplier)) %>%
  summarize(variance = sum(p_win * (multiplier - exp_ret)^2), 
            std_dev = sqrt(variance), 
            exp_ret = min(exp_ret)) %>%
  mutate(coef_of_var = std_dev/exp_ret) %>%
  left_join(luckybit_addrIDs_raw, by = "Name")

luckybit_addrIDs
```


Calculate measures for all identified users:

- daily_n_avg: number of bets placed per days active
- days_played: distict calendar days with placed bets
- playspan: number of days between first and last active calendar date

- n_games: number of bets placed
- avg_bet: mean value of placed bets
- avg_log_bet: mean value of placed bets on a log10 scale
- max_bet: maximum of bet placed
- max_cv: maximum of coefficient of vairance of games chosen
- avg_cv: average coefficient of variance chosen over games chosen
- log_avg_cv: avg_cv but weighted with logarithmic bet size
- total_bet: total ammount of bet placed

```{r}
daily_avg <- bets_complete %>%
  mutate(date = as.Date(as.POSIXct(block_timestamp, origin="1970-01-01"))) %>%
  group_by(userID, date) %>%
  summarize(daily_n = n()) %>%
  group_by(userID) %>%
  summarize(daily_n_avg = sum(daily_n)/n(),
            days_played = n(),
            playspan = abs(as.integer(difftime(max(date), min(date), units = "days")))+1
            )


gamer_stats <- bets_complete %>%
  left_join(luckybit_addrIDs %>% select(addrID, game_name = Name, coef_of_var), by = "addrID") %>%
  group_by(userID) %>%
  summarize(
    n_games = n(),
    avg_bet = mean(value),
    avg_log_bet = mean(log10(value)),
    max_bet = max(value),
    max_cv = max(coef_of_var),
    avg_cv = sum(coef_of_var)/n(),
    log_avg_cv = sum(log10(value) * coef_of_var)/sum(log10(value)),
    total_bet = sum(value)
  ) %>%
  left_join(daily_avg, by = "userID")

gamer_stats
```

# Visual inspection of measures over the population

## Number of games

```{r}
gamer_stats %>%
  ggplot(aes(x = n_games)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +  # Histogram
  scale_x_log10() +  # Logarithmic transformation on x-axis
  labs(x = "Number of bets per user (log 10 scale)",
       y = "Frequency") +
  theme_minimal()
```

## Mean logarithmic bet per user

```{r}
gamer_stats %>%
  ggplot(aes(x = avg_log_bet)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +  # Histogram
  #scale_x_log10() +  # Logarithmic transformation on x-axis
  labs(x = "mean log bet per user",
       y = "Frequency") +
  theme_minimal()
```

## Median bet per user (log 10 scale)

```{r}
gamer_stats %>%
  ggplot(aes(x = max_bet)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +  # Histogram
  scale_x_log10() +  # Logarithmic transformation on x-axis
  labs(x = "Median bet per user (log 10 scale)",
       y = "Frequency") +
  theme_minimal()
```

## Maximum risk taken per user

```{r}
gamer_stats %>%
  ggplot(aes(x = max_cv)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +  # Histogram
  #scale_x_log10() +  # Logarithmic transformation on x-axis
  labs(x = "Max risk",
       y = "Frequency") +
  theme_minimal()
```

## Average risk taken per user

```{r}
gamer_stats %>%
  ggplot(aes(x = avg_cv)) +
  geom_histogram(fill = "blue", color = "black") +  # Histogram
  scale_x_log10() +  # Logarithmic transformation on x-axis
  labs(x = "avg risk",
       y = "Frequency") +
  theme_minimal()
```

## Total bet per user

```{r}
gamer_stats %>%
  ggplot(aes(x = total_bet)) +
  geom_histogram(fill = "blue", color = "black") +  # Histogram
  scale_x_log10() +  # Logarithmic transformation on x-axis
  labs(title = "Histogram with Logarithmic X-Axis",
       x = "total bet",
       y = "Frequency") +
  theme_minimal()
```

## Daily bets average per user

```{r}
gamer_stats %>%
  ggplot(aes(x = daily_n_avg)) +
  geom_histogram(fill = "blue", color = "black") +  # Histogram
  scale_x_log10() +  # Logarithmic transformation on x-axis
  labs(title = "Histogram with Logarithmic X-Axis",
       x = "daily bets average",
       y = "Frequency") +
  theme_minimal()
```

## Number of days active per player

```{r}
gamer_stats %>%
  ggplot(aes(x = days_played)) +
  geom_histogram(fill = "blue", color = "black") +  # Histogram
  scale_x_log10() +  # Logarithmic transformation on x-axis
  labs(title = "Histogram with Logarithmic X-Axis",
       x = "days played",
       y = "Frequency") +
  theme_minimal()
```

## Playspan per player

```{r}
gamer_stats %>%
  ggplot(aes(x = playspan)) +
  geom_histogram(fill = "blue", color = "black") +  # Histogram
  scale_x_log10() +  # Logarithmic transformation on x-axis
  labs(title = "Histogram with Logarithmic X-Axis",
       x = "days between start and finish",
       y = "Frequency") +
  theme_minimal()
```

# Clustering

```{r}
scaled_gamer_stats <- gamer_stats %>%
  mutate(
    log_max_bet = log10(max_bet),
    log_total_bet = log10(total_bet)
  ) %>%
  select(n_games, 
         days_played, 
         avg_cv, 
         log_max_bet, 
         log_total_bet) %>%
  scale()

set.seed(123) #fix seed
trim_parameter = .01 #set trimming parameter to 1%

km.pred <- trimcluster::trimkmeans(scaled_gamer_stats, 2, trim = trim_parameter, runs = 100)

#check clustering descriptors

scaled_gamer_stats_2 <- scaled_gamer_stats %>% 
  bind_cols(cluster = km.pred$classification) %>%
  filter(cluster != 3) %>%
  select(-cluster)

classific <- km.pred$classification[km.pred$classification != 3]
```

## Cluster validation statistics

```{r}
fpc::cluster.stats(d = dist(scaled_gamer_stats_2), classific)
```

## Median statistics of clusters

Results used in Table 2.

```{r}
zero_label <- rep(0,length(km.pred$classification))

gamer_stats_clustered <- gamer_stats %>% # add the unclustered sample as cluster "zero"
  cbind(cluster = km.pred$classification) %>%
  rbind(gamer_stats %>%
         cbind(cluster =zero_label))

aggregate(gamer_stats_clustered, by=list(cluster=gamer_stats_clustered$cluster), median)
```

## IQR statistics of clusters

Results used in Table 2.

```{r}

aggregate(gamer_stats_clustered, by=list(cluster=gamer_stats_clustered$cluster), IQR, na.rm = T)
```

# Save clustering

```{r}
gamer_stats %>%
  select(userID) %>%
  cbind(cluster = km.pred$classification) %>%
  write.csv(".\\data\\clustered_users.csv")
```

